---
title: "Synthetic Data Generation: SDSS 2022 Short Course"
author: "Joshua Snoke"
date: '2022-06-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(synthpop)

set.seed(405323)
```

## I-CEM 1901 Scottish Census Data Example

For this code, we will use a portion of the 1901 Scottish Census. Thanks to the SYLLS and the synthpop team for preparing and sharing this file, which can be found at: https://synthpop.org.uk/resources.html.

NOTE: synthpop does not work well with the tidyverse, so I will be using traditional R instead.

```{r ICEM data}
load('i-cem.rdata')
## let's just work with a subset of the variables
icem_data = data[, c('sex',
                     'age',
                     'mar_stat',
                     'disability',
                     'employ',
                     'ctry_bth',
                     'hsize',
                     'totrooms',
                     'pperroom')]
rm(data)

head(icem_data,
     n = 10)
```

## Simple Synthesis Model

First, let's produce a simple synthetic data set for just a few demographic variables in the data.

```{r simple synthesis}

simple_icem = icem_data[icem_data$mar_stat != 'Divorced', 
                        c('sex',
                          'age',
                          'mar_stat',
                          'employ')]
## Here is the syntax
synthetic_icem = syn(simple_icem,
                     ## Here we will tell it to use sequential parametric models
                     method = 'parametric')

## A few outputs of interest
synthetic_icem$method
synthetic_icem$visit.sequence
synthetic_icem$predictor.matrix

## We can change the visit.sequence (order of synthesized variables)
synthetic_icem_2 = syn(simple_icem,
                       method = 'parametric',
                       ## Here we change the order of synthesis
                       visit.sequence = c('employ', 'age', 'mar_stat', 'sex'))

## A few outputs of interest
synthetic_icem_2$method
synthetic_icem_2$visit.sequence
synthetic_icem_2$predictor.matrix

## We can use non-parametric CART models instead of GLMs 
synthetic_icem_3 = syn(simple_icem,
                       ## Here we will tell it to use sequential CART models
                       method = 'cart')

## A few outputs of interest
synthetic_icem_3$method
synthetic_icem_3$visit.sequence
synthetic_icem_3$predictor.matrix

```

## Utility Evaluation

```{r utility measures}

## We can compare the marginal distributions between the confidential data and synthetic data
compare(synthetic_icem, simple_icem)

## We can also get visuals of multivariate distributions
multi.compare(synthetic_icem, simple_icem, var = "employ", by = c("sex", "mar_stat"))

## If we want to compare the fit of a linear model, we can use the following function
## outcome is whether individual is an employer (rather than a worker or unemployed)
icem_fit = glm(I(employ == 'E') ~ ., data = simple_icem, family = "binomial")  
summary(icem_fit)


# Now with synthetic data
## just main effects
synthetic_icem_fit = glm.synds(I(employ == 'E') ~ ., data = synthetic_icem, family = "binomial")  

## we can get the estimated coefficients and standard errors for the synthetic model
summary(synthetic_icem_fit)


## we can compare the models visually and based on their point estimates and confidence intervals
## note that the confidential data, not the confidential glm object, is provided
compare(synthetic_icem_fit, simple_icem)


```


## Some Other Useful Synthesis Options

Lastly, we cover a couple common options that you may use when synthesizing data. First, creating new methods for variables that are constructed based on other variables. Second, increasing the number of levels of a factor variable.

```{r other options}

## we first create an object to say the type of synthesis for each variable
## default is CART for all variables
synthetic_method = rep('cart', ncol(icem_data)) 
## it needs to be named with the corresponding variable names
names(synthetic_method) = colnames(icem_data)
## but we change pperroom to be just hsize divided by totrooms for consistency in the data
synthetic_method[names(synthetic_method) == "pperroom"] <- "~I(hsize[!is.na(totrooms)] / totrooms[!is.na(totrooms)])"


## country of birth has a lot of levels, so we set the maxfaclevels to be larger enough
## note that this can slow down the synthesis, especially if using parametric models
synthetic_expanded_icem = syn(icem_data,
                              ## using our predefined methods object
                              method = synthetic_method,
                              maxfaclevels = 60)

synthetic_expanded_icem$method

## we can check that the synthesis rule worked as it should have
summary(synthetic_expanded_icem$syn$hsize / synthetic_expanded_icem$syn$totrooms - 
          synthetic_expanded_icem$syn$pperroom)

## we can look at the results for this expanded synthesis
compare(synthetic_expanded_icem, icem_data)

```

## Try One or More of the Following Yourself

1. Compare the synthetic data to the confidential data using the 'multi.compare' function or compare the fit of GLM models using the synthetic data and confidential data.

2. The following variables combine to make the hsize: 
  - nfamgteq15 
  - nfamlt15
  - nboarders
  - nlodgers
  - nvisitors
  - nkn
  - nservants
  
  Produce a synthetic data sets that synthesizes these variables and then creates the hsize variable, so that it is consistent.

3. The variables occlab1, occlab2, occlab3 are nested occupation labels. Synthesis them using the nested method (for example for occlab2 nested within occlab1 use the method 'nested.occlab1')

4. Or use your own data!


