---
title: "Variable Suppression and Recoding: SDSS 2022 Short Course"
author: "Joshua Snoke"
date: '2022-06-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
™options(dplyr.summarise.inform = FALSE)
```

## Simulated Data Example

Start by loading in a simulated data set of a teacher survey. The file contains individual and school characteristics. Print out the first 25 rows to get an idea of the variables.

```{r survey data}
simulated_survey_data = read_csv('simulated_suppression_data.csv',
                                 show_col_types = FALSE) %>%
  mutate_all(as.factor)
print(simulated_survey_data,
      n = 25)
```

We can also tabulate the variables to get an idea of the distributions.

```{r data summary}
summary(simulated_survey_data,
        maxsum = 8)
```

## Identification Risk Among the Sample

First, let's look just at frequencies of characteristics among those in the survey. We have a lot of characteristics, so the majority of teachers are unique if we used every variable and level as provided.

```{r survey freq trivial}
trivial_freq = simulated_survey_data %>%
  ## here we are computing all combinations of the characteristics
  group_by_all() %>%
  summarize(freq = n())

table(trivial_freq$freq)

## We can also plot the frequencies
ggplot(trivial_freq, aes(x = freq)) + 
  geom_histogram(binwidth = 1) +
  theme_bw()

## We can calculate risk based on 1 / Freq
summary(1 / trivial_freq$freq)

```

State is a pretty identifying characteristic, so let's see what happens when we suppress the state variable.

```{r survey freq no state}
no_state_freq = simulated_survey_data %>%
  ## here we are dropping state from our calculations (i.e., suppressing it)
  group_by_at(vars(-State)) %>%
  summarize(freq = n())

ggplot(no_state_freq, aes(x = freq)) + 
  geom_histogram(binwidth = 1) +
  theme_bw()

## how many teachers are still unique?
sum(no_state_freq$freq == 1)

## Risk has dropped but is still high
summary(1 / no_state_freq$freq)


```

We want to do more, so let's collapse some of the small levels for Race, Experience, and School Size

```{r survey freq collapsed}
no_state_collapsed_freq = simulated_survey_data %>%
  ## here we collapse levels for Race and Experience
  mutate(Race = recode(Race,
                       `AIAN` = 'Other',
                       `NHPI` = 'Other',
                       `Asian` = 'Other',
                       `Multirace` = 'Other',
                       `Black` = 'Other',
                       `Hispanic` = 'Other'),
         Experience = recode(Experience,
                             `Lessthan3` = 'Lessthan10',
                             `3to9` = 'Lessthan10'),
         School_Size = recode(School_Size,
                              `Lessthan250` = 'Lessthan500',
                              `250to499` = 'Lessthan500')) %>%
  ## here we still drop state from our calculations (i.e., suppressing it)
  group_by_at(vars(-State)) %>%
  summarize(freq = n())

ggplot(no_state_collapsed_freq, aes(x = freq)) + 
  geom_histogram(binwidth = 1) +
  theme_bw()

## how many teachers are still unique?
sum(no_state_collapsed_freq$freq == 1)

## Risk has dropped again but is still high
summary(1 / no_state_collapsed_freq$freq)


```


## Identification Risk with Population Frequencies

Now, let's assume that participation in the survey is not known. In this case, we need to factor in the likelihood that an individual was sampled into the survey. We do this using population frequencies.

```{r population data}
simulated_population = read_csv('simulated_suppression_population.csv',
                                show_col_types = FALSE) %>%
  mutate_at(vars(-pop_total), as.factor)

```

We calculate risk using both the survey frequencies and the population frequencies.

```{r population freq trivial}

## survey freq are the same as before
trivial_pop_freq = simulated_survey_data %>%
  ## here we are computing all combinations of the characteristics
  group_by_all() %>%
  summarize(freq = n()) %>%
  ## merge with population freq
  left_join(simulated_population) %>%
  ## Risk is now based on the population frequencies
  mutate(risk = 1 / pop_total,
         sample_unique = freq == 1)

## Risk is now the multiplication of the two
summary(trivial_pop_freq$risk)
## Much lower than before, but still some population uniques
sum(trivial_pop_freq$pop_total == 1)

```

If we drop state and Race in this case, our risk is substantially lower.

```{r population freq no state}

## survey freq are the same as before
no_state_pop_freq = simulated_survey_data %>%
  ## here we are computing all combinations of the characteristics
  group_by_at(vars(-State, 
                   -Race)) %>%
  summarize(freq = n()) %>%
  ## merge with population freq that are calculated based on dropped variables
  left_join(simulated_population %>%
              group_by_at(vars(-State, 
                               -Race, 
                               -pop_total)) %>%
              summarize(pop_total = sum(pop_total))) %>%
  ## Risk is now based on the population frequencies
  mutate(risk = 1 / pop_total,
         sample_unique = freq == 1)

summary(no_state_pop_freq$risk)
## No longer any population uniques
sum(no_state_pop_freq$pop_total == 1)

## we can visual risk values against the population totals
ggplot(no_state_pop_freq, aes(y = risk, x = pop_total)) + 
  geom_point(size = 2) +
  theme_bw()

```

## Try One or More of the Following Yourself

1. Produce a combination of variables/levels with max risk less than 0.1.

2. Produce a combination of variables/levels with mean risk less than 0.01

3. Produce a combination of variables/levels with total risk less than 1.

4. Or use your own data!




